---
title: "Homework1"
author: "Xueying Liu"
date: "8/28/2020"
output: pdf_document
---

# Problem 2
## Part A

```{r}
library(knitr)
knitr::kable(c("R","Python","SAS"),col.names = "3 Things I Want to Learn")
```

## Part B
Bernoulli distribution(p):$$P(X=x|p)=p^x(1-p)^{1-x}; \;\; x=0,1;\;0<=p<=1.$$
Binomial distribution(n,p):$$P(X=x|n,p)={n \choose x}p^x(1-p)^{n-x}; \;\;x=0,1,2,...,n; \;0<=p<=1.$$
Poisson distribution($\lambda$):$$P(X=x|\lambda)=\frac{e^{-\lambda}\lambda^x}{x!}; \;\;x=0,1,...; \;0<=\lambda<\infty.$$

# Problem 3
## Steps in performing reproducible research:
$Step1$: For every result, keep track of how it was produced, record every involved steps.
Challenges:Sometimes we just try some possible methods, manually recording each step can take a lot of works.

$Step2$: Archive the exact versions of all external programs used.
Challenges:Remember to note the exact names and versions of the main programs you use.

$Step3$: Use version control system to track evolution of codes and help the exact reproduction of results.

$Step4$: Record all intermediate result,  randomness and raw data behinds plots. Connect textual statements to underlying results.
Challenges: It may be hard to locate the exact result underlying and supporting the statement from a large pool of different analyses with various versions.

$Step5$: Provide public access to your codes and results.
Challenges: Be prepared to respond to any requests for further data or methodology details by peers.

# Problem 4
We will use the cars dataset in the R and plot a scatter plot and a histogram of it.
```{r}
library(ggplot2)
ggplot(data = cars,mapping = aes(x=dist, y=speed))+
  geom_point(color="blue")
ggplot(data=cars,aes(x=dist))+
  geom_histogram(color="black",fill="white", bins=50)
```

