---
title: "HW2-906338789"
author: "Xueying Liu"
date: "9/10/2020"
output: pdf_document
---
# Problem 3
I think I will definitely use it since I am always revising my codes with with new idea, so it is useful when I make a mistake and want to revert back and campare the old version with the new one. It can also be helpful by sharing it with other people so that people can work together on the same problem.

# Problem 4
## a. Sensory data from five operators
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(DataCombine)
library(tidyverse)
library(knitr)
```

```{r sensory_import,cache=FALSE,warning=FALSE}
## getting "https://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/Sensory.dat"
url="https://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/Sensory.dat"
sensory_data_raw=fread(url,header = TRUE,fill=TRUE,skip="Item",data.table = FALSE)
saveRDS(sensory_data_raw,"sensory_data_raw.RDS")
sensory_data_raw=readRDS("sensory_data_raw.RDS")
## filling the first column with Item number
for(i in 0:9)
{
  sensory_data_raw[(3*i+2):(3*i+3),]=c(i+1,sensory_data_raw[(3*i+2):(3*i+3),])
  }
```

To tidy the data, we need to push operator into a column.

```{r Sensory_tidy_baseR,cache=FALSE}
sensory_data_tidy=data.frame(rep(sensory_data_raw$Item,5),stack(sensory_data_raw[,-1]))
colnames(sensory_data_tidy)=c("Item","value","Operator")
head(sensory_data_tidy)
```
We have converted the dataframes to tidy data frames using the base function. Here is a summary and boxplot of the data:
```{r sensory_summary_baseR, include=TRUE,echo=FALSE}
knitr::kable(summary(sensory_data_tidy))
boxplot(sensory_data_tidy$value~sensory_data_tidy$Operator)
```


Then we choose to use $tidyverse()$ function to tidy the raw data:
```{r sensory_tidy_tidyverse,cache=FALSE}
# stack and fix column names using tidyverse
sensory_data_tv=sensory_data_raw %>%
                     gather(key="operator",value="value",2:6)
head(sensory_data_tv)

```



## b. Gold Medal performance for Olympic Men’s Long Jump
```{r gold_import,cache=FALSE,warning=FALSE,message=FALSE}
## getting "https://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/LongJumpData.dat"
url="https://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/LongJumpData.dat"
gold_data_raw=fread(url,data.table = FALSE)
saveRDS(gold_data_raw,"gold_data_raw.RDS")
gold_data_raw=readRDS("gold_data_raw.RDS")
gold_data_raw=gold_data_raw[,1:8]
colnames(gold_data_raw)=c("Year","LongJump","Year","LongJump","Year","LongJump","Year","LongJump")
```

```{r gold_tidy_baseR,cache=FALSE,warning=FALSE,message=FALSE}
gold_data_tidy=data.frame(rbind(gold_data_raw[,1:2],gold_data_raw[,3:4]
                                ,gold_data_raw[,5:6],gold_data_raw[,7:8]))
## Drop the raws with missing value
gold_data_tidy=DropNA(gold_data_tidy)
head(gold_data_tidy)
```
We have converted the dataframes to tidy data frames using the base function. Here is a summary and plot of the data:
```{r gold_summary_baseR, include=TRUE,echo=FALSE，fig.height=3.6}
knitr::kable(summary(gold_data_tidy))
plot(gold_data_tidy$LongJump~gold_data_tidy$Year)
```
Then we choose to use $tidyverse()$ function to tidy the raw data:
```{r include=FALSE}
colnames(gold_data_raw)=c("Year1","LongJump1","Year2","LongJump2","Year3","LongJump3","Year4","LongJump4")
```

```{r gold_tidy_tidyverse,cache=FALSE}
# stack and fix column names using tidyverse
gold_data_tv=data.frame(gather(gold_data_raw,key = "year",value="year",1,3,5,7)[,6],
                        gather(gold_data_raw,key="LongJump",value="LongJump",2,4,6,8)[,6])
colnames(gold_data_tv)=c("Year","LongJump")
head(gold_data_tv)

```



## c. Brain weight (g) and body weight (kg) for 62 species
```{r brain_import,warning=FALSE,message=FALSE}
## getting "https://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/BrainandBodyWeight.dat"
url="https://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/BrainandBodyWeight.dat"
brain_data_raw=fread(url,data.table = FALSE,header =TRUE,fill=TRUE)
saveRDS(brain_data_raw,"brain_data_raw.RDS")
brain_data_raw=readRDS("brain_data_raw.RDS")
colnames(brain_data_raw)=c(rep(c("Body_Wt","Brain_Wt"),3))
```

```{r brain_tidy_baseR, warning=FALSE,message=FALSE}
brain_data_rbind=DropNA(data.frame(rbind(brain_data_raw[,1:2],
                                        brain_data_raw[,3:4],brain_data_raw[,5:6])))
brain_data_tidy=data.frame(stack(brain_data_rbind))
colnames(brain_data_tidy)=c("weight","Brain/Body")
head(brain_data_tidy)
```
We have converted the dataframes to tidy data frames using the base function. Here is a summary of the data:
```{r brain_summary_baseR, include=TRUE,echo=FALSE}
knitr::kable(summary(brain_data_tidy))
boxplot(brain_data_tidy$weight~brain_data_tidy$`Brain/Body`)
```

Then we choose to use $tidyverse()$ function to tidy the raw data:
```{r brain_tidy_tidyverse,cache=FALSE}
# stack and fix column names using tidyverse
brain_data_tv=gather(brain_data_rbind,key="Brain/Body",value="value",Body_Wt:Brain_Wt)
head(brain_data_tv)

```

## d. Triplicate measurements of tomato yield for two varieties of tomatos at three planting densities
```{r tomato_import,warning=FALSE,message=FALSE}
## getting "https://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/tomato.dat"
url="https://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/tomato.dat"
tomato_data_raw=fread(url,data.table = FALSE,skip="1000")
saveRDS(tomato_data_raw,"tomato_data_raw.RDS")
tomato_data_raw=readRDS("tomato_data_raw.RDS")
```
```{r tomato_tidy_baseR, warning=FALSE,message=FALSE}

tomato_data=data.frame(varity=rep(tomato_data_raw[1:2,1],3),stack(tomato_data_raw[,-1]))
colnames(tomato_data)=c("variety","value","planting_density")
value=do.call("rbind", strsplit(tomato_data$value, ","))
value=data.frame(apply(value,2,as.numeric))
colnames(value) = c("value1","value2","value3")
tomato_data_split=data.frame(cbind(tomato_data$variety,value,tomato_data$planting_density))
tomato_data_tidy=data.frame(variety=rep(tomato_data_split[,1],3),stack(tomato_data_split),
                             plant_density=rep(tomato_data_split[,5],3))[,-3]
head(tomato_data_tidy)
```
We have converted the dataframes to tidy data frames using the base function. Here is a summary and boxplots of the data:
```{r tomato_summary_baseR, include=TRUE,echo=FALSE}
knitr::kable(summary(tomato_data_tidy))
boxplot(tomato_data_tidy$values~tomato_data_tidy$variety)
boxplot(tomato_data_tidy$values~tomato_data_tidy$plant_density)
```

Then we choose to use $tidyverse()$ function to tidy the raw data:
```{r tomato_tidy_tidyverse,cache=FALSE}
# stack and fix column names using tidyverse
tomato_data_tv=gather(tomato_data_split,key="value",value="value",value1:value3)[,-3]
head(tomato_data_tv)

```
