---
title: "HW5_Liu"
author: "Xueying Liu"
date: "10/27/2020"
output: pdf_document
---

```{r setup, include=FALSE}
library(knitr)
library(sur)
library(ggplot2)
library(dplyr)
library(ggpubr)
knitr::opts_chunk$set(echo = TRUE,warning = FALSE, message = FALSE)
```

# Problem 3
```{r echo=FALSE}
library(readr)
EdStatsData <- read_csv("C:/Users/yxm/Desktop/5014 Stat Package/HW5/EdStatsData.csv")
```
```{r}
dim(EdStatsData)
```

There are 886930 data in total in the complete data set.
```{r}
# clean the data
# We will only consider the data till 2020
cleaned <- EdStatsData[,1:53]

# remove the row that is NA in all the year from 1970 to 2020
cleaned <- cleaned[apply(cleaned[,5:53],1,function(x)any(!is.na(x))),]
dim(cleaned)
```
After we removed the rows that the data is missing from 1970 to 2020, there are 354575 data remains.

```{r}
## choose 2 countries
country <- rbind(cleaned[cleaned$`Country Code`=="EAP",], cleaned[cleaned$`Country Code`=="ARB",]) 
country$`Indicator Code`<- factor(country$`Indicator Code`)

## Choose useful indicator to compare
enrolment <- apply(country[country$`Indicator Name`=="Adjusted net enrolment rate, primary, both sexes (%)",5:53],1,mean,na.rm=TRUE)
population <- apply(country[country$`Indicator Name`=="Population, total",5:53],1,mean,na.rm=TRUE)
unemployment <- apply(country[country$`Indicator Name`=="Unemployment, total (% of total labor force)",5:53],1,mean,na.rm=TRUE)
literacy <- apply(country[country$`Indicator Name`=="Adult literacy rate, population 15+ years, both sexes (%)",5:53],1,mean,na.rm=TRUE)
primary <- apply(country[country$`Indicator Name`=="Enrolment in primary education, both sexes (number)",5:53],1,mean,na.rm=TRUE)
secondary <- apply(country[country$`Indicator Name`=="Enrolment in secondary education, both sexes (number)",5:53],1,mean,na.rm=TRUE)
tertiary <- apply(country[country$`Indicator Name`=="Enrolment in tertiary education, all programmes, both sexes (number)",5:53],1,mean,na.rm=TRUE)
GDP <- apply(country[country$`Indicator Name`=="GDP per capita (current US$)",5:53],1,mean,na.rm=TRUE)
GNI <- apply(country[country$`Indicator Name`=="GNI (current US$)",5:53],1,mean,na.rm=TRUE)
labor <- apply(country[country$`Indicator Name`=="Labor force, total",5:53],1,mean,na.rm=TRUE)
populationgrowth <- apply(country[country$`Indicator Name`=="Population growth (annual %)",5:53],1,mean,na.rm=TRUE)
table <- format(rbind(population,populationgrowth,GDP,GNI,enrolment,unemployment,labor,literacy,
                          primary,secondary,tertiary),scientific = FALSE,digits = 2)
```
```{r echo=FALSE}
colnames(table) <- c("East Asia & Pacific","Arab World")
row.names(table) <- c("Population","Population growth (annual %)","GDP","GNI","Adjusted net enrolment rate","Unemployment, total (% of total labor force)","Labor force, total","Adult literacy rate, population 15+ years, both sexes (%)","Enrolment in primary education, both sexes (number)","Enrolment in secondary education, both sexes (number)","Enrolment in tertiary education, all programmes, both sexes (number)")
kable(table)
```

This is the summary table of the average of these indicators over 30 years from 1970 to 2020 in East Asia & Pacific area and Arab World area.

# Problem 4
In this problem, we are going to model the linear relationship between the GDP per capita in 2010 and the GDP per capita in 2000 and 1990.
```{r}
## Only explore GDP indecator in each country
data.4 <- cleaned[cleaned$`Indicator Name`=="GDP per capita (current US$)",]

## select independent variable 1990 and 2000 and dependent variable 2010
data.4 <- data.4[complete.cases(data.4[ , c('1990','2000','2010')]),]

## Do the linear regression
fit <- lm(data.4$`2010`~data.4$`1990`+data.4$`2000`)
```
Create plots for this linear regression model:
```{r }
par(mfrow=c(2,3),oma = c(0, 0, 2, 0))
plot(predict(fit),residuals(fit),xlab = "Predicted Value",ylab = "Residual")
abline(h=0)


plot(predict(fit),rstudent(fit),xlab = "Predicted Value",ylab = "RStudent")
abline(h=-3)
abline(h=3)

plot(leverage(fit),rstudent(fit),xlab = "Leverage",ylab = "Rstudent")
abline(h=-3)
abline(h=3)
abline(v=0.3)

qqnorm(rstudent(fit),xlab = "Quantile",ylab = "Residual", main = "")
qqline(rstudent(fit))

plot(predict(fit),data.4$`2010`,xlab = "Predicted Value",ylab = "GDP")

cooks <- cooks.distance(fit)
plot(cooks,xlab = "Observation",ylab = "Cook's D")
mtext("Fit Diagnostics of Linear Model",outer = TRUE)
```

# Problem 5
We can recreate the plot in problem 3 using ggplot2 functions.
```{r}
fig1 <- qplot(x=predict(fit),y=residuals(fit))+xlab("Predicted Value")+ylab("Residual")+
  geom_hline(yintercept =0)

fig2 <- qplot(x=predict(fit),y=rstudent(fit))+xlab("Predicted Value")+ylab("Rstudent")+
  geom_hline(yintercept =3,linetype="dotted")+geom_hline(yintercept =-3,linetype="dotted")

fig3 <- qplot(x=leverage(fit),y=rstudent(fit))+xlab("Leverage")+ylab("Rstudent")+
  geom_hline(yintercept =3,linetype="dotted")+geom_hline(yintercept=-3,linetype="dotted")+
  geom_vline(xintercept =0.3,linetype="dotted")

fig4 <- ggplot()+geom_qq(aes(sample=residuals(fit)))+geom_qq_line(aes(sample=residuals(fit)))+
  xlab("Quantile")+ylab("Residual")
  
fig5 <- qplot(x=predict(fit),y=data.4$`2010`)+xlab("Predicted Value")+ylab("GDP")

fig6 <- qplot(y=cooks)+xlab("Observation")+ylab("Cook's D")

figlist <- list(fig1,fig2,fig3,fig4,fig5,fig6)
fitdiag <- ggarrange(plotlist = figlist, nrow = 2, ncol = ceiling(length(figlist)/2))+ggtitle("Fit Diagnostics of Linear Model")
annotate_figure(fitdiag,
               top = text_grob("Fit Diagnostics of Linear Model", face = "bold", size = 18))

```


